\bookmark[dest=\HyperLocalCurrentHref, level=1]{Conclusion}
\section{Conclusion} \label{sec6}
In this paper, we have presented a comprehensive survey of non-line-of-sight (NLOS) imaging, covering the foundational physical models and algorithms established through 2021 and the significant advances made from 2022 to 2026. The field has matured substantially during this period, progressing from laboratory demonstrations with minute-long acquisition times to real-time video of dynamic room-sized scenes, and from a primarily optical inverse problem to a broader computational sensing paradigm spanning light, RF, sound, terahertz waves, and robotic perception.

\textbf{Progress on fundamental challenges.}
The three core challenges identified in the 2021 survey---ill-posedness with low SNR, limited resolution, and long data acquisition time---have all seen meaningful progress. On SNR, photon-efficient imaging~\cite{liuPhotonEfficientNLOS2022} and first-photon event stamping~\cite{liFastFirstPhoton2022} have demonstrated high-quality reconstruction at sub-photon-per-pixel levels. Resolution has been pushed to $\sim0.6\,\text{mm}$ via active wavefront shaping (UNCOVER~\cite{caohighresolutionnlos2022}), while picosecond-resolution up-conversion detection has shown that sub-millimeter axial resolution can also be pursued through extreme timing precision. Most dramatically, real-time 4\,fps NLOS video of room-sized dynamic scenes has been demonstrated~\cite{yeRealTimeNLOS2024}, and scan-free systems operating at up to 19\,fps have been reported~\cite{zhangRealTimeScanFree2024}, fundamentally changing the prospects for practical NLOS deployment.

\textbf{Deep learning advances.}
The deep learning component of NLOS imaging has seen a transformation in architecture diversity and sophistication. Beyond the U-Net and ResNet baselines of the 2021 era, the field has adopted transformers (NLOST~\cite{liNLOST2023}, TransiT~\cite{liTransiT2025}), Mamba/state-space models for video~\cite{liMambaTemporalConsistency2024}, graph neural networks~\cite{suDGNLOS2025}, and diffusion model priors~\cite{suModelGuidedDiffusion2024}. Neural implicit surface representations (NLOS-NeuS~\cite{fujimuraNLOSNeuS2023}) have extended beyond voxel volumes to enable smooth 3D surface reconstruction. Recent works further broaden this trajectory: \href{https://arxiv.org/abs/2510.09205}{Time-Resolved Transformer} studies photon-efficient LOS/NLOS transient 3D reconstruction with spatio-temporal attention; \href{https://arxiv.org/abs/2506.08470}{MARMOT} introduces masked autoencoder pretraining for transient/NLOS data; and \href{https://arxiv.org/abs/2508.09655}{NANO} uses a noise-adapted neural operator to improve sparse and photon-starved reconstruction. Learnable physical priors~\cite{sunGeneralizable2025}, phasor-field enhancement under partial acquisition~\cite{choLEAP2024}, and domain reduction strategies~\cite{shimDomainReduction2024} have improved generalization, while unsupervised methods~\cite{cuiVirtualScanning2024} and untrained network priors~\cite{wuUntrainedDecoder2022} have reduced dependence on large training datasets.

\textbf{Forward models, rendering, and reconstruction infrastructure.}
A parallel trend is the improvement of the computational infrastructure around NLOS imaging. Dedicated NLOS rendering tools~\cite{royoNLOSTransientRendering2022,plackFastDifferentiable2023} and real-world datasets such as NLOS-Track~\cite{wangPropagateCalibrate2023} have better supported training and evaluation. Newer tools such as \href{https://arxiv.org/abs/2510.25660}{mitransient} extend this direction by bringing time-resolved, polarized, and differentiable transient light transport into Mitsuba 3, including NLOS scene setups and capture-noise modeling. On the reconstruction side, \href{https://arxiv.org/abs/2308.02782}{structure-sparsity regularization} improves DLCT-style inversion under low-SNR conditions, while \href{https://arxiv.org/abs/2508.02003}{Quasi-Fresnel Transform} and optimized FFT/NUFFT sampling strategies~\cite{sultanOptimizedSamplingNLOS2025} explore lighter-weight and more flexible transient inversion. Sparse and irregular active acquisition has also become a central modeling issue: arbitrary illumination/detection reconstruction~\cite{liuArbitraryPatternNLOS2023}, few-shot signal-surface collaborative regularization~\cite{liuFewShotSSCR2022}, and LEAP~\cite{choLEAP2024} show how Bayesian virtual-confocal formulations, mixed-dimensional priors, and phasor-domain measurement completion can preserve reconstruction quality when the relay samples are incomplete, irregular, noisy, or very sparse. Ptychographic NLOS~\cite{songPtychographicNLOS2024} adds a coherent wavefront-retrieval route by using relay-wall modulation and ptychographic diversity for depth-resolved hidden-object visualization. Iterating the transient light transport matrix~\cite{sultanIteratingTLTM2024} pushes this model-centric trajectory further by treating a measured relay-surface TLTM as a first-order system that can be computationally focused into the hidden scene to obtain second-order transport, hidden-scene relighting, and dual-photography-style measurements. Complementing these method papers, the recent comparative ToF study by Marco~\etal~\cite{marcoComprehensiveToFNLOS2026} is valuable because it places representative NLOS inverse models under a common forward-model and hardware-aware evaluation framework, clarifying which limitations arise from physics, capture geometry, photon counts, and algorithm-specific assumptions.

\textbf{Passive NLOS advances.}
Passive NLOS imaging has progressed from simple 2D reconstruction with occluders to real-time passive tracking of moving persons (PAC-Net~\cite{wangPropagateCalibrate2023}) and full-color 3D passive NLOS imaging with ordinary cameras~\cite{czajkowskiTwoEdge2024}. Hyperspectral and multispectral fusion (Hyper-NLOS~\cite{chenHyperNLOS2024}) and the exploitation of rough surfaces as relay surfaces~\cite{liRoughSurfaces2025} have broadened the environmental conditions and surface types compatible with passive NLOS sensing. Event-camera passive NLOS, represented by \href{https://arxiv.org/abs/2209.13300}{NLOS-ES} and \href{https://arxiv.org/abs/2404.05977}{event-enhanced passive NLOS with physical embedding}, shows that neuromorphic sensing can preserve fast motion cues that are difficult to capture with conventional frame cameras.

\textbf{New modalities and scenes.}
Beyond optical NLOS, mmWave radar NLOS reconstruction (HoloRadar~\cite{laiHoloRadar2025}) and ultrasound synthetic aperture NLOS imaging~\cite{ultrasoundNLOS2025} have demonstrated full 3D scene reconstruction in non-optical modalities, enabling operation in darkness, through walls, and under conditions where optical methods fail. Optical NLOS itself is also moving beyond clean free-space propagation: scattering-media phasor-field analysis~\cite{luesiaScatteringPhasor2023} studies how virtual-wave NLOS behaves when the hidden scene and relay path are affected by volumetric scattering. The latest frontier extends even further: \href{https://arxiv.org/abs/2605.29098}{GeRaF 2.0} studies RF/radar 3D reconstruction through boxes; \href{https://arxiv.org/abs/2602.11471}{RIS-assisted radar} actively controls the RF propagation environment for around-corner sensing, while dual-beam RIS radar~\cite{yasmeenDualBeamRIS2026} studies the practical effect of one-bit quantized RIS beam control and reciprocal two-way scattering; passive non-reconfigurable electromagnetic skins~\cite{tornielliEMSkinsNLOS2024} provide a lower-complexity metasurface route for multi-view near-field NLOS radar imaging; \href{https://arxiv.org/abs/2603.13981}{distributed MIMO/ISAC NLOS imaging} uses NLOS multipath for joint synchronization and off-grid imaging; \href{https://arxiv.org/abs/2505.08240}{N$^2$LoS} explores tag-assisted mmWave NLOS localization; and \href{https://arxiv.org/abs/2205.05066}{terahertz NLOS} uses lossy-mirror environmental reflections for around-obstacle imaging. Acoustic NLOS has also diversified from relay-surface reconstruction to \href{https://arxiv.org/abs/2506.08471}{relay-free passive acoustic localization} via edge diffraction. In robotics, single-photon-LiDAR navigation work~\cite{youngNavigationNLOS2024} and \href{https://arxiv.org/abs/2510.10506}{SuperEx} connect NLOS measurements with indoor exploration and mapping, demonstrating that partial hidden-space information can be valuable even when it is used primarily for decision making rather than final-quality reconstruction.

\textbf{Geometry, relay surfaces, and representation.}
The classical assumption of a planar diffuse relay wall is increasingly being relaxed. \href{https://arxiv.org/abs/2606.21270}{3D Gaussian Transient Rendering} models arbitrary relay surface geometries with differentiable transient rendering, while ptychographic NLOS~\cite{songPtychographicNLOS2024}, arbitrary illumination and detection patterns~\cite{liuArbitraryPatternNLOS2023}, scattering-media NLOS~\cite{luesiaScatteringPhasor2023}, and active corner-camera systems expand the range of feasible acquisition layouts. In the RF regime, engineered reflectors such as RISs, Van-Atta-array IRSs, and static electromagnetic skins show that the propagation environment itself can become part of the sensing hardware. These developments suggest that future NLOS systems will be less tied to laboratory walls and more capable of exploiting real environmental geometry.

\textbf{Remaining challenges and future directions.}
Despite remarkable progress, several important challenges remain open. (1) \textit{Outdoor and strong ambient light operation:} most existing systems require controlled or low-ambient-light conditions. Coherent sensing approaches~\cite{huangCoherentNLOS2024} and temporal gating designs~\cite{riccardoFastGatedSPAD2022} are promising, but robust outdoor operation at long range remains difficult. (2) \textit{Large-scale real training data:} the gap between synthetic and real-world NLOS measurements continues to limit generalization. Automated data collection systems, more realistic rendering, and domain adaptation methods are needed. (3) \textit{Non-Lambertian and complex relay surfaces:} most algorithms still assume diffuse relay walls; exploiting specular~\cite{liRoughSurfaces2025}, non-planar surfaces~\cite{guNonPlanar2023}, arbitrary relay geometries, and volumetric scattering media in a unified framework remains an open problem. (4) \textit{End-to-end system integration:} bridging the gap between research demonstrations and practical deployed systems---handling calibration, robustness, and computational cost---is a critical next step. (5) \textit{Semantic NLOS understanding:} going beyond shape reconstruction to semantic tasks such as \href{https://arxiv.org/abs/2003.14414}{human pose estimation}, action recognition, and object identification in hidden scenes represents an important emerging frontier.

In summary, NLOS imaging has transitioned from a laboratory curiosity to a technology on the verge of practical deployment across multiple modalities and application domains. Continued progress on the challenges outlined above will be essential for realizing the transformative potential of seeing around corners.
